---
title: Maintain your brain, Primary Cognitive Composite Variable
subtitle: CACE analysis
output: html_document
params:
  data : "engagement"
  response : "cogmeanz_full"
  predictors : !r c("rc_trial_time","group_type_id")
  covariates : !r c("myb_age", "myb_gender","cl_edu_yrs","rcal_modules_sum","adri_baseline")
  id : "id_mask"
  final : true
---

## `r params$response`

- Primary Cognitive Composite Variable
    - Cogstate Detection
    - Cogstate Identification
    - Cogstate One Back
    - Cambridge Brain Sciences Spatial (tokens) Search
    - Cambridge Brain Sciences Grammatical Reasoning
    - Cogstate One Card Learning
    - Cambridge Brain Sciences Paired Associates

These analyses use `bl_adritotal` (as named in the original SPSS file) as the ADRI baseline covariate.

These results exclude those participants who were eligible for fewer than 2 modules but who were randomised. This results in N=6104.

This is the analysis for the complier-averaged causal effect.  In this analysis, randomisation group is used as an instrumental variable.

- Define "complience", which can be thought of as whether the treatment was received (and done by the participant). 
  - In the intervention group, we will define this as completing at least 60% of content for all eligible modules.
  - In the control group, they are all assume to have 0 treatment received no treatment (`tx_received`=0)
  
- Two stage process:
  - Regress `tx_received` on randomisation group and save probabilities.
  - Predict outcome from probabilities
  - Estimate the effect (this is the CACE estimate)
  - Bootstrap CIs 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```


```{r read-data, message=FALSE, warning=FALSE}
# read in functions and data
source("functions.R")

myb_dat <- read.csv(paste0("../../Data/",params$data,".csv"))

```

Inference for differences in `r params$predictors[1]`  trend by `r params$predictors[2]` for  `r params$response`

## Data summary


```{r outcome-summary, message=!params$final, warning=!params$final}
myb_dat %>% 
  group_by(.dots = lapply(params$predictors, as.symbol)) %>% 
  summarize_at(.vars = params$response, .funs = list(Mean = mean, Sd = sd, N_obs = ~length(.)), na.rm=T)
```

## Defining Compliance groups

```{r}

# myb_dat <- myb_dat %>%
#   mutate(tx_received = case_when(
#     group_type_id == 1 ~ 0,
#     group_type_id == 2 ~ overall_engagement_bin,
#   ))

myb_dat <- myb_dat %>%
  mutate(tx_received = ifelse(group_type_id == 1, 0, overall_engagement_bin))
   
```

# Predict compliance from group (for baseline only)

```{r}
b <- myb_dat %>%
  filter(rc_trial_time==1)

# control group should have mean, min and max of 0,
# int group should have mean>0 and variability
b %>% 
  group_by(group_type_id) %>% 
  summarize(Mean = mean(tx_received),
            SD=sd(tx_received),
            Min=min(tx_received),
            Max=max(tx_received))


compliance <- glm(tx_received ~ group_type_id, data=b, family = binomial)
# get predicted values
b$pred_compliance <- predict(compliance, type = "response")
# get participant IDs
  b <- b %>%
    dplyr::select(pred_compliance,id_mask)
  # merge back with full
myb_dat_wc <- myb_dat %>%
  left_join(b, by="id_mask")
```


```{r}

# CACE estimate
# run the model, using time*predicted compliance as the interaction.
# should the emmeans be calculated at 0 and 1 (engagement score) or 
# 0 and 2?
# in looking at the 0/1 estimates, I think this is correct.


unadj_mod <- lmer(cogmeanz_full ~ factor(rc_trial_time)*pred_compliance + 
                                  (1|id_mask), data=myb_dat_wc)
# summary(unadj_mod)
emm<- emmeans(unadj_mod, ~rc_trial_time*pred_compliance,
        at = list(pred_compliance=c(0,1)) )

did<-contrast(emm, list(contr = c(1,0,0,-1,-1,0,0,1)))
confint(did)$estimate

plot(unadj_mod)


```

## Bootstrap

https://stackoverflow.com/questions/11919808/block-bootstrap-from-subject-list

```{r}

library(boot)
# myfit <- function(x, i) {
#   mydata <- do.call("rbind", lapply(i, function(n) subset(myb_dat, id_mask==x[n])))
#   b <- mydata %>%
#     filter(rc_trial_time==1)
# 
# compliance <- glm(tx_received ~ group_type_id, data=b, family = binomial)
# 
# # get predicted values
#   b$pred_compliance <- predict(compliance, type = "response")
# # get participant IDs
#   b <- b %>%
#     dplyr::select(pred_compliance,id_mask)
#   # merge back with full
# mydata <- mydata %>%
#   left_join(b, by="id_mask")
# 
# 
# unadj_mod <- lmer(cogmeanz_full ~ factor(rc_trial_time)*pred_compliance +
#                                   (1|id_mask), data=mydata)
# emm<- emmeans(unadj_mod, ~rc_trial_time*pred_compliance,
#         at = list(pred_compliance=c(0,1)) )
# 
# did<-contrast(emm, list(contr = c(1,0,0,-1,-1,0,0,1)))
# summary(did)$estimate
# 
# }
# 
# id_masks <- unique(myb_dat$id_mask)
# 
# start_time <- Sys.time()
# b1 <- boot(id_masks, myfit, 999)
# Sys.time() - start_time
# 
# save(b1, file = "CACE_boot_cat")
load("CACE_boot_cat")



CI <- boot.ci(b1, type = "perc")
pval <- mean(c(confint(did)$estimate,b1$t)<0)
res <-  data.frame(response = params$response,
            adjusted = "No",
            base_3yr_did = round(confint(did)$estimate,3),
            LCL = round(CI$percent[4],3),
            UCL = round(CI$percent[5],3),
            `p.value` = ifelse (pval < 0.001, "<0.001",paste("= ",round(pval,3)))
)

res

write.csv(res,file = paste0("../../Results/baseline_sd/",params$response,"_CACE_engagement_.csv"), row.names = FALSE)


evidence_unadj = ifelse(res$p.value[1]<0.05, "evidence", "no evidence")

```




## Result

We classified someone as "engaged" if they competed at least 60% of units within every module for which they were eligible. 

The association of the engagement on outcomes was analysed using a complier average causal effect (CACE) analysis with randomisation group as instrumental variable (Dunn 2005) using a two-step approach with block bootstrap percentile confidence intervals. The first step predicts engagement from randomisation group with a logistic regression. The second step is a linear mixed model with with time, predicted engagement score and their interaction as predictors. 

The unadjusted difference in difference value from baseline to three year follow up was `r res$base_3yr_did[1]` ( 95%CI: `r res$LCL[1]`- `r res$UCL[1]`, p`r res$p.value[1]`) baseline standard deviations. 




Analysis by Nancy Briggs and Gordana Popovic on `r Sys.Date()`

## Reference

Dunn G, Maracy M, Tomenson B. Estimating treatment effects from randomized clinical trials with noncompliance and loss to follow-up: the role of instrumental variable methods. Stat Methods Med Res. 2005 Aug;14(4):369-95. doi: 10.1191/0962280205sm403oa. PMID: 16178138.
