---
title: "Creating some data for writing MI programs"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)



library(tidyverse)
library(kableExtra)
```

Updated "`r Sys.Date()`"


Nov 11, 2022 update


  - As per email, Nov 5, 2022 from Henry Brodaty:
    - ADRI at baseline (`adri_baseline` for the covariates) should now be based on `adritotal` (`bl_adritotal` in the original SPSS file)
    - Removed completely from the datasets the participants eligible for <=1 module
      - this results in N=6104.

Previous Updates:

- As per email from Megan, ADRI variable is `sum_newadri_bl`. If this variable is missing, then `adti_total` is used
- new variable to use as covariate is called `adri_baseline`



Getting some data with the individual test items for writing MI program code.

This includes the following changes to the raw data:  

- Reversing coding reaction time tests: Detection (test_id==5), Identification (test_id==6),
One Back (test_id==7)


## Importing data

All datasets were exported to CSV format from SPSS.  The CSV files are imported into R.

```{r dataimport}

# Status Index data
st.index <- read.csv("../Data/mybm_status_index_20220422.csv",header = TRUE,
                 fileEncoding = "UTF-8-BOM")

# Demographics data
demo <- read.csv("../Data/mybm_b_demos_v3.1_20181018.csv",header = TRUE,
                 fileEncoding = "UTF-8-BOM")

# Group Allocation
group <- read.csv("../Data/mybm_b_group_v1.0_20181018.csv",header = TRUE,
                 fileEncoding = "UTF-8-BOM")

# Trial Activities
st.trial <- read.csv("../Data/mybm_status_trial_20220422.csv",header = TRUE,
                 fileEncoding = "UTF-8-BOM")

# Cognition Outcomes
cogout <- read.csv("../Data/mybm_full_cogout_20220422.csv",header = TRUE,
                 fileEncoding = "UTF-8-BOM")


# adritotal and some potential imputation predictors
adri <- read.csv("../Data/tempmh_mybm_missing_20220617.csv",header = TRUE,
                 fileEncoding = "UTF-8-BOM")
```


## Relevant variables for merging and analysis


- st.index
  - randomisation_status
  - id_mask
  - myb_trial_status
  - rcal_modules_sum
- st.trial
  - activity_title
  - module_title
  - rc_title
  - tpa_status
- cogout
  - activity_title
  - resultPrimaryOutcome
  - test_name
  - test_id
- demo
  - myb_age
  - myb_gender
  - cl_edu_yrs
- adri
  - adritotal
  - sum_newadri_bl
  
  
## Merging process

- Create `BaselineSample` dataset:
  - `st.index`: select rc_randomisation_status==1
    - Keep relevant variables: id_mask, rc_randomisation_status, myb_trial_status, rcal_modules_sum
  - Merge with `demo`
    - Keep relevant variables: myb_age, myb_gender, cl_edu_yrs
  - Merge with `group`
    - Keep relevant variables: group_type_id
  - merge with `adri`
    - Keep relevant variables: adritotal, sum_newadri_bl (for analysis covariates)
    - Keep variables for potential inclusion in imputation model: CalcK10SCore, rcal_mybBMI, rcal_total_energy, PREDIMEDTotal, smoking, everchol, everbp, LLS, CalcAnxietyScore, rc_phq9_total, rc_wellb_status, rc_paqu_status, rc_medh_status, rc_leq_status, rc_diet_status, rc_adri_status, rc_myb1_status

- Create `cogprimary` dataset:
  - Merge `BaselineSample` with  `st.trial`
    - Keep relevant variables: rc_trial_time, myb_trial_status, activity_title, module_title, rc_title, tpa_status
    - Keep observations (test scores) tpa_status=="Completed"
  - Merge with `cogout` (key variables are id_mask, module_title(st.trial)/ module_name(cogout), activity_title(st.trial)/activity_name(cogout) )
    - Keep relevant variables: test_name,test_id, resultPrimaryOutcome
  
  
  

```{r}

BaselineSample <- st.index %>%
  filter(rc_randomisation_status==1) %>%
  dplyr::select(id_mask,rc_randomisation_status,myb_trial_status, rcal_modules_sum)  %>%  
  left_join(., group, by="id_mask") %>%
  relocate(group_type_id, .after=id_mask) %>%
  left_join(.,demo, by="id_mask") %>%
  dplyr::select(id_mask,rc_randomisation_status, myb_trial_status,
                rcal_modules_sum, group_type_id, myb_age, myb_gender,
                cl_edu_yrs) %>%
  left_join(., select(adri, -c(rc_randomisation_status)), by="id_mask") %>%
  mutate(adri_baseline=adritotal) %>%
  dplyr::select(id_mask,rc_randomisation_status, myb_trial_status,
                rcal_modules_sum, group_type_id, myb_age, myb_gender,
                cl_edu_yrs, adritotal, sum_newadri_bl, adri_baseline, CalcK10Score, 
                rcal_mybBMI, rcal_total_energy, PREDIMEDTotal,
                smoking, everchol, everbp, LLS, CalcAnxietyScore, rc_phq9_total, rc_wellb_status,
                rc_paqu_status, rc_medh_status, rc_leq_status, rc_diet_status, rc_adri_status,
                rc_myb1_status) 




# Select subset of people
BaselineSample_randsample <- BaselineSample %>%
  sample_frac(.2) %>%
  dplyr::select(id_mask)



# Full dataset for imputation
cogprimary <- BaselineSample %>%
  left_join(., st.trial, by=c("id_mask","rc_randomisation_status")) %>%
  dplyr::select(id_mask,rc_randomisation_status,rc_trial_time, myb_trial_status, 
                rcal_modules_sum, activity_title,module_title,rc_title,tpa_status, group_type_id, myb_age, myb_gender,
                cl_edu_yrs, adritotal, sum_newadri_bl, adri_baseline, CalcK10Score, 
                rcal_mybBMI, rcal_total_energy, PREDIMEDTotal,
                smoking, everchol, everbp, LLS, CalcAnxietyScore, rc_phq9_total, rc_wellb_status,
                rc_paqu_status, rc_medh_status, rc_leq_status, rc_diet_status, rc_adri_status,
                rc_myb1_status) %>%
  filter(tpa_status=="Completed") %>%
  left_join(.,cogout, by=c("id_mask","rc_randomisation_status", "module_title"="module_name",
                           "activity_title"="activity_name" ) ) %>%
  filter(activity_title=="Cognitive Assessment (Thinking)" | 
        activity_title=="Cognitive Assessment (Cards)") %>%
  dplyr::select(id_mask,rc_randomisation_status,rc_trial_time, myb_trial_status,
                rcal_modules_sum,
                activity_title,module_title,rc_title,tpa_status, test_name,test_id,
                resultPrimaryOutcome, group_type_id, myb_age, myb_gender,
                cl_edu_yrs, adritotal, sum_newadri_bl, adri_baseline, CalcK10Score, 
                rcal_mybBMI, rcal_total_energy, PREDIMEDTotal,
                smoking, everchol, everbp, LLS, CalcAnxietyScore, rc_phq9_total, rc_wellb_status,
                rc_paqu_status, rc_medh_status, rc_leq_status, rc_diet_status, rc_adri_status,
                rc_myb1_status) 
          # DO NOT NEED for IMPUTATION
          # %>%
          # mutate(ed_group=ifelse(cl_edu_yrs<12,0,1)) # For the zscore calculation


MI_testing <- cogprimary %>%
  
  # reverse coding Detection, Identification, One Back
  mutate(resultPrimaryOutcome=ifelse(test_id %in% c(5,6,7), 
                                     (-1 * resultPrimaryOutcome), 
                                     resultPrimaryOutcome)) %>%


       # create domain category variable
       mutate(domain=ifelse(test_id %in% c(5,6), 1,
                     ifelse(test_id %in% c(7, 21,24), 2,
                     ifelse(test_id %in% c(8,23), 3, 0))) )  %>%
      
        # Must make sure that each person at each timepoint has at least 2 scores
        # First, count valid scores at each time for each person within each domain
        # count number of valid test scores for each domain within each person
        # group_by(id_mask, rc_trial_time, domain) %>%
        # mutate(n_tests=n())  %>%
        # group_by(rc_trial_time, test_id) %>%
        # filter out people with only 1 score for a domain
        # if you want to know how many are filtered out, you can create a new dataset here.
        #filter(n_tests>=2) %>%

  filter(rc_trial_time %in% c(1,10,11,12) ) 
```


Variables to use for imputation model:

Analysis model variables

  * group_type_id 
  * rc_trial_time
  * group_type_id * rc_trial_time
  * myb_age
  * myb_gender
  * cl_edu_yrs
  * rcal_modules_sum
  * adri_baseline

Extra variables for code testing

  * Online modules: rcal_eligibility_bts, rcal_eligibility_pom, rcal_eligibility_pa, rcal_eligibility_diet
      * These are binary yes/ no variables
  * rc_race: race
  * rc_preflang: preferred language
  * rc_leqses: SES
  * Various baseline scores and status indicators: 
    * CalcK10Score
    * rcal_mybBMI
    * rcal_total_energy
    * PREDIMEDTotal,
    * smoking
    * everchol
    * everbp
    * LLS
    * CalcAnxietyScore
    * rc_phq9_total
    * rc_wellb_status
    * rc_paqu_status
    * rc_medh_status
    * rc_leq_status
    * rc_diet_status
    * rc_adri_status
    * rc_myb1_status


```{r imputionvars}

## add in some more  variables for imputation  model

# Module elgibility & Demographics
MI_vars <- st.index %>%
  filter(rc_randomisation_status==1) %>%
  dplyr::select(id_mask,rcal_eligibility_bts, rcal_eligibility_pom, rcal_eligibility_pa,
                rcal_eligibility_diet)  %>%
  left_join(., demo, by="id_mask") %>%
  select(id_mask,rcal_eligibility_bts, rcal_eligibility_pom, rcal_eligibility_pa,
                rcal_eligibility_diet, rc_race, rc_preflang, rc_leqses )



MI_testing <- MI_testing %>%
  left_join(., MI_vars, by="id_mask")

```


  

```{r}
write.csv(MI_testing,"../MI/MI_testing_fullSample.csv", row.names = TRUE)
```

Now, use the 20% random Baseline sample for testing.


```{r}

MI_testing_randSample <- MI_testing %>%
  right_join(., BaselineSample_randsample, by="id_mask")

write.csv(MI_testing_randSample,"../MI/MI_testing_randSample.csv", row.names = TRUE)
```
