---
title: Maintain your brain, Executive Function
output: html_document
params:
  data : "cogprimary"
  response : "domainmean_full_2"
  predictors : !r c("rc_trial_time","group_type_id")
  covariates : !r c("myb_age", "myb_gender","cl_edu_yrs","rcal_modules_sum","adri_baseline")
  id : "id_mask"
  final : true
---

## `r params$response`

- Domain 2: Executive Function
    - Cogstate One Back
    - Cambridge Brain Sciences Spatial (tokens) Search
    - Cambridge Brain Sciences Grammatical Reasoning

These analyses use `bl_adritotal` (as named in the original SPSS file) as the ADRI baseline covariate.

These results exclude those participants who were eligible for fewer than 2 modules but who were randomised. This results in N=6104.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r read-data, message=FALSE, warning=FALSE}
# read in functions and data
source("functions.R")
myb_dat <- read.csv(paste0("../../Data/",params$data,".csv"))
```

Inference for differences in `r params$predictors[1]`  trend by `r params$predictors[2]` for  `r params$response`



## Data summary



```{r outcome-summary, message=!params$final, warning=!params$final}
myb_dat %>% 
  group_by(.dots = lapply(params$predictors, as.symbol)) %>% 
  summarize_at(.vars = params$response, .funs = list(Mean = mean, Sd = sd, N_obs = ~length(.)), na.rm=T)
```



```{r fit-models, message=!params$final, warning=!params$final}

# if want to run
if(!params$final){
  
  # run models
  fitted_mods <- MYB_fit(params$response, params$predictors, 
                       params$covariates, params$id, myb_dat, times = plannedtimes)
  # save models
  save(fitted_mods,file = paste0("../../Results/baseline_sd/",params$response))
  
} else {
  load(file = paste0("../../Results/baseline_sd/",params$response))
}


```



```{r plot-residuals, message=!params$final, warning=!params$final}
#plot models

if(!params$final){
  lapply(fitted_mods, plot, form = cond)
}
```

```{r print-summary, message=!params$final, warning=!params$final}
#check summary is okay

if(!params$final){
  lapply(fitted_mods, summary)
}
```

## Results

```{r inference, message=!params$final, warning=!params$final}

if(!params$final){
  
  # calculate CI
  res_tab = MYB_inf(fitted_mods, params$response)
  
  # save models
  write.csv(res_tab,file = paste0("../../Results/baseline_sd/",params$response,"_table.csv"), row.names = FALSE)
  

  
} else {
  res_tab <- read.csv(file = paste0("../../Results/baseline_sd/",params$response,"_table.csv"))
  
  
}

res <- res_tab %>% 
  transmute(response,
            adjusted,
            base_3yr_did = round(planned,3),
            LCL = round(asymp.LCL,3),
            UCL = round(asymp.UCL,3),
            zstat,
            pz,
            Chisq,
            `p.value` = ifelse (`p.value` < 0.001, "<0.001",paste("= ",round(`p.value`,3)))
)

res


evidence_unadj = ifelse(res_tab$p.value[1]<0.05, "evidence", "no evidence")
evidence_adj = ifelse(res_tab$p.value[2]<0.05, "evidence", "no evidence")

```

```{r, message=!params$final, warning=!params$final}

# calculate all ems by time and group
group_time_ems <- emmeans(fitted_mods$unadj_mod,  ~ group_type_id + rc_trial_time)

# contrast to baseline by group
comp_to_baseline <- contrast(group_time_ems, rc_trial_time , "trt.vs.ctrl", by=  "group_type_id")

time_diff <- c(1:3)

to_bl_plot_df <- cbind(summary(comp_to_baseline)[,1:3], 
                       confint(comp_to_baseline)[,6:7], 
                       `Visit (Years)` = rep(c(1:3), 2))
# add baseline 0
to_bl_plot_df[7:8,] <- NA
to_bl_plot_df$group_type_id[7:8] = c(1,2)
to_bl_plot_df$estimate[7:8] = c(0,0)
to_bl_plot_df$`Visit (Years)`[7:8] = 0


to_bl_plot_df %>% 
  ggplot(aes(`Visit (Years)`, estimate,color=factor(group_type_id), group=factor(group_type_id))) + 
  geom_point() + 
  geom_line()+
  geom_linerange(aes(ymin=asymp.LCL, ymax=asymp.UCL), size=2, alpha=.5) + 
  theme_classic() + 
  ylab("Mean change from baseline") +
  scale_color_discrete(name="Randomisation Group")  + 
   theme(legend.position = "bottom") 

ggsave(file = paste0(params$response,"_",params$data,"_plot.png"), width = 12, height = 8, units = "cm")



to_bl_plot_df %>% 
  transmute(group_type_id, `Visit (Years)`,
         estimate,
         LCL = asymp.LCL, UCL = asymp.UCL) %>% 
  arrange(`Visit (Years)`,group_type_id) %>% 
write.csv(file = paste0(params$response,"_",params$data,"_plotdata.csv"), row.names = FALSE)


```


## Result

Covariates included in adjusted models: 

  - `myb_age`: Age at Baseline 
  - `myb_gender`: Reported gender
  - `cl_edu_yrs`: Years of education
  - `rcal_modules_sum`: Number of modules person was eligible for
  - `adri_baseline`: ADRI at Baseline

A linear mixed model with a random effect for patient and fixed effects of time point, randomisation group and their interaction was run. An additional analysis including covariates of `r paste(params$covariates, collapse=", ")` was run.  We report estimates of the group difference in the change from baseline to three years follow up with 95% confidence interval, and associated Z statistic and p-value. The units of the effect size are standard deviations at baseline. 

The unadjusted difference between randomisation groups of change from baseline to three year follow up was `r res$base_3yr_did[1]` ( 95%CI: `r res$LCL[1]`- `r res$UCL[1]`,Z=`r res$zstat[1]`, p`r res$pz[1]`) baseline standard deviations. 

The difference between randomisation groups of change from baseline to three year follow up was `r res$base_3yr_did[2]` ( 95%CI: `r res$LCL[2]` - `r res$UCL[2]`,Z=`r res$zstat[2]`,p`r res$pz[2]`) baseline standard deviations, adjusting for `r paste(params$covariates, collapse=", ")`.


Analysis by Nancy Briggs and Gordana Popovic on `r Sys.Date()`


