---
title: 'Secondary Outcome: Raw ADRI'
author: "Nancy Briggs & Gordana Popovic"
date: "`r Sys.Date()`"
output: html_document
params:
  data : "adri_time"
  response : "rawadri"
  predictors : !r c("rc_trial_time","group_type_id")
  covariates : !r c("myb_age", "myb_gender","cl_edu_yrs")
  id : "id_mask"
  final : true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(stringsAsFactors=FALSE)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

library(tidyverse)
library(kableExtra)
```

## `r params$response`


Raw ADRI does not have any missing values at baseline. N=6104

```{r read-data, message=FALSE, warning=FALSE}
# read in functions and data
source("functions.R")
myb_dat <- read.csv(paste0("../../Data/",params$data,".csv"))

```




## Data summary

confirming Ns. 

```{r outcome-summary, message=!params$final, warning=!params$final}
# commented out because the N function is not excluding missing observations
# myb_dat %>% 
#   group_by(.dots = lapply(params$predictors, as.symbol)) %>% 
#   summarize_at(.vars = params$response, .funs = list(Mean = mean, Sd = sd, N_obs = ~length(.)), na.rm=T)


myb_dat %>%
  group_by(rc_trial_time) %>%
  summarise(cn = sum(!is.na(rawadri))) 

myb_dat %>%
  group_by(group_type_id,rc_trial_time) %>%
  summarise(
            cn = sum(!is.na(rawadri)), cmean=mean(rawadri, na.rm=T), 
            csd=sd(cleanadri, na.rm=T))




```

```{r fit-models, message=!params$final, warning=!params$final}

# if want to run
if(!params$final){
  
  # run models
  fitted_mods <- MYB_fit(params$response, params$predictors, 
                       params$covariates, params$id, myb_dat, times = plannedtimes)
  # save models
  save(fitted_mods,file = paste0("../../Results/Prelim/",params$response))
  
} else {
  load(file = paste0("../../Results/Prelim/",params$response))
}


```



```{r plot-residuals, message=!params$final, warning=!params$final}
#plot models

if(!params$final){
  lapply(fitted_mods, plot, form = cond)
}
```

```{r print-summary, message=!params$final, warning=!params$final}
#check summary is okay

if(!params$final){
  lapply(fitted_mods, summary)
}
```


### Results

```{r inference, message=!params$final, warning=!params$final}

if(!params$final){
  
  # calculate CI
  res_tab = MYB_inf(fitted_mods, params$response)
  
  # save models
  write.csv(res_tab,file = paste0("../../Results/Prelim/",params$response,"_table.csv"), row.names = FALSE)
  

  
} else {
  res_tab <- read.csv(file = paste0("../../Results/Prelim/",params$response,"_table.csv"))
  
  
}
```

```{r}
res <- res_tab %>% 
  transmute(response,
            adjusted,
            base_3yr_did = round(planned,3),
            LCL = round(asymp.LCL,3),
            UCL = round(asymp.UCL,3),
            zstat,
            pz,
            Chisq,
            `p.value` = ifelse (`p.value` < 0.001, "<0.001",paste("= ",round(`p.value`,3)))
)

res


evidence_unadj = ifelse(res_tab$p.value[1]<0.05, "evidence", "no evidence")
evidence_adj = ifelse(res_tab$p.value[2]<0.05, "evidence", "no evidence")

```


```{r, message=!params$final, warning=!params$final}

# calculate all ems by time and group
group_time_ems <- emmeans(fitted_mods$unadj_mod,  ~ group_type_id + rc_trial_time)

# contrast to baseline by group
comp_to_baseline <- contrast(group_time_ems, rc_trial_time , "trt.vs.ctrl", by=  "group_type_id")

time_diff <- c(1:3)

to_bl_plot_df <- cbind(summary(comp_to_baseline)[,1:3], 
                       confint(comp_to_baseline)[,6:7], 
                       `Visit (Years)` = rep(c(1:3), 2))
# add baseline 0
to_bl_plot_df[7:8,] <- NA
to_bl_plot_df$group_type_id[7:8] = c(1,2)
to_bl_plot_df$estimate[7:8] = c(0,0)
to_bl_plot_df$`Visit (Years)`[7:8] = 0


to_bl_plot_df %>% 
  ggplot(aes(`Visit (Years)`, estimate,color=factor(group_type_id), group=factor(group_type_id))) + 
  geom_point() + 
  geom_line()+
  geom_linerange(aes(ymin=asymp.LCL, ymax=asymp.UCL), size=2, alpha=.5) + 
  theme_classic() + 
  ylab("Mean change from baseline") +
  scale_color_discrete(name="Randomisation Group")  + 
   theme(legend.position = "bottom") 

ggsave(file = paste0(params$response,"_",params$data,"_plot.png"), width = 12, height = 8, units = "cm")



to_bl_plot_df %>% 
  transmute(group_type_id, `Visit (Years)`,
         estimate,
         LCL = asymp.LCL, UCL = asymp.UCL) %>% 
  arrange(`Visit (Years)`,group_type_id) %>% 
write.csv(file = paste0(params$response,"_",params$data,"_plotdata.csv"), row.names = FALSE)


```


```{r, message=F, warning=F, include=F}
emmip(fitted_mods$unadj_mod,  group_type_id ~ rc_trial_time, CIs = TRUE) + 
  theme_classic() + 
  ylab(params$response)
```



## Result

Covariates included in adjusted models: 

  - `myb_age`: Age at Baseline 
  - `myb_gender`: Reported gender
  - `cl_edu_yrs`: Years of education
  
There are a total of `r fitted_mods$adj_mod@Gp[2]` participants included in the mixed model analysis.  

A linear mixed model with a random effect for patient and fixed effects of time point, randomisation group and their interaction was run. An additional analysis including covariates of `r paste(params$covariates, collapse=", ")` was run. **NOTE: baseline ADRI was not included as a covariate in these analyses.**

We report estimates of the group difference in the change from baseline to three years follow up with 95% confidence interval, and associated Z statistic and p-value. **Note: ADRI score is modelled without baseline standardisation, units of effect size are in the units of ADRI score.**


The unadjusted difference between randomisation groups of change from baseline to three year follow up was `r res$base_3yr_did[1]` ( 95%CI: `r res$LCL[1]`- `r res$UCL[1]`,Z=`r res$zstat[1]`, p`r res$pz[1]`) . 

The difference between randomisation groups of change from baseline to three year follow up was `r res$base_3yr_did[2]` ( 95%CI: `r res$LCL[2]` - `r res$UCL[2]`,Z=`r res$zstat[2]`,p`r res$pz[2]`) , adjusting for `r paste(params$covariates, collapse=", ")`.



Analysis by Nancy Briggs and Gordana Popovic on `r Sys.Date()`



From SAP:  Unadjusted and adjusted analyses.
A linear mixed model with fixed effects of treatment group, time and the interaction will be performed. A random effect for individual will be specified. Unadjusted and adjusted estimates of the group difference at 3 years along with p-values and 95% CIs will be reported. Adjusted analyses will include the same variables as indicated in the primary outcome analysis.