---
title: Maintain your brain, Minutes Aerobic Exercise
output: html_document
params:
  data : "pa"
  response : "ncal_msaerobic_wk"
  predictors : !r c("rc_trial_time","group_type_id")
  covariates : !r c("myb_age", "myb_gender","cl_edu_yrs","rcal_modules_sum","adri_baseline")
  id : "id_mask"
  final : false
---

## `r params$response`

- Secondary Outcome: Minutes Aerobic Exercise

These analyses use `bl_adritotal` (as named in the original SPSS file) as the ADRI baseline covariate.

These results exclude those participants who were eligible for fewer than 2 modules but who were randomised. This results in N=6104.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)


```


```{r read-data, message=FALSE, warning=FALSE}
# read in functions and data
source("functions.R")
myb_dat <- read.csv(paste0("../../Data/",params$data,".csv"))  %>%
    mutate(rc_trial_time= as.factor(rc_trial_time))

library(glmmTMB)
library(DHARMa)

```

Inference for differences in `r params$predictors[1]`  trend by `r params$predictors[2]` for  `r params$response`

## Data summary

A 5-number summary is presented for the values here as the data are highly skewed.

```{r outcome-summary, message=!params$final, warning=!params$final}

p <- c(0, 0.25, 0.5, 0.75, 1.00)
p_names <- map_chr(p, ~paste0(.x*100))
p_funs <- map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>% 
  set_names(nm = p_names)

# Percentiles
myb_dat %>%
  group_by(rc_trial_time, group_type_id) %>%
  summarize_at(vars(params$response), funs(!!!p_funs)) 

# Means
myb_dat %>% 
  group_by(.dots = lapply(params$predictors, as.symbol)) %>% 
  summarize_at(.vars = params$response, .funs = list(Mean = mean, Sd = sd, N_obs = ~length(.)), na.rm=T)

```

### Histograms

```{r,hist, message=!params$final, warning=!params$final }


myb_dat %>% ggplot(aes(x=ncal_msaerobic_wk)) + geom_histogram() +
  facet_grid(group_type_id~rc_trial_time) + theme_classic() 



```

```{r fit-models, message=!params$final, warning=!params$final}

# if want to run
if(!params$final){
  # if(FALSE){
  # run models

  
  unadj_mod <- glmmTMB(ncal_msaerobic_wk ~   rc_trial_time*group_type_id + (1 | id_mask) , 
                       data=myb_dat, family = tweedie )
  unadj_null <- glmmTMB(ncal_msaerobic_wk ~ rc_trial_time + group_type_id + (1 | id_mask) ,
                        data=myb_dat, family = tweedie )
  adj_mod <- glmmTMB(ncal_msaerobic_wk ~ myb_age + myb_gender + cl_edu_yrs + rcal_modules_sum +  
                       adri_baseline + rc_trial_time * group_type_id + (1 | id_mask) ,
                     data=myb_dat, family = tweedie )
  adj_null <- glmmTMB(ncal_msaerobic_wk ~ myb_age + myb_gender + cl_edu_yrs + rcal_modules_sum +  
                        adri_baseline + rc_trial_time + group_type_id + (1 | id_mask) , 
                      data=myb_dat, family = tweedie )
  
  # save models
  save(unadj_mod,file = paste0("../../Results/baseline_sd/",params$response,"_unadj_mod"))
  save(unadj_null,file = paste0("../../Results/baseline_sd/",params$response,"_unadj_null"))
  save(adj_mod,file = paste0("../../Results/baseline_sd/",params$response,"_adj_mod"))
  save(adj_null,file = paste0("../../Results/baseline_sd/",params$response,"_adj_null"))
  
} else {
  load(file = paste0("../../Results/baseline_sd/",params$response,"_unadj_mod"))
  load(file = paste0("../../Results/baseline_sd/",params$response,"_unadj_null"))
  load(file = paste0("../../Results/baseline_sd/",params$response,"_adj_mod"))
  load(file = paste0("../../Results/baseline_sd/",params$response,"_adj_null"))
  }

fitted_mods = list(adj_mod = adj_mod,
              unadj_null = unadj_null,
              unadj_mod = unadj_mod,
              adj_null = adj_null)

```



```{r plot-residuals, message=!params$final, warning=!params$final}
#plot models


if(!params$final){
  # if(FALSE){
  # lapply(fitted_mods, plot, form = cond)

  # sim= simulateResiduals(unadj_mod, n = 5000)
  # resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
  # scatter.smooth(resid~predict(unadj_mod, re.form = NA,type="response"),
  #                lpars =list(col = "red", lwd = 3, lty = 3))  
  # 
  # sim= simulateResiduals(adj_mod, n = 5000)
  # resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
  # scatter.smooth(resid~predict(adj_mod, re.form = NA,type="response"),
  #                lpars =list(col = "red", lwd = 3, lty = 3))
  # 
  # sim= simulateResiduals(unadj_null, n = 5000)
  # resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
  # scatter.smooth(resid~predict(unadj_null, re.form = NA,type="response"),
  #                lpars =list(col = "red", lwd = 3, lty = 3))
  # 
  # sim= simulateResiduals(adj_null, n = 5000)
  # resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
  # scatter.smooth(resid~predict(adj_null, re.form = NA,type="response"),
  #                lpars =list(col = "red", lwd = 3, lty = 3))  

 
  }
```

```{r print-summary, message=!params$final, warning=!params$final}
#check summary is okay

if(!params$final){
  summary(unadj_null)
  summary(unadj_mod)
  summary(adj_null)
  summary(adj_mod)
}
```

## Results

```{r inference, message=!params$final, warning=!params$final}


if(!params$final){
  
  
  #unadjusted
  anova_call <- anova(unadj_null,unadj_mod)
  likrat <-  data.frame(Chisq = paste0(round(anova_call$Chisq[2],3),"(",
                                       anova_call$Df[2],")"),
                        pvalue = anova_call$`Pr(>Chisq)`[2])
  em <- emmeans(unadj_mod,  ~ rc_trial_time*group_type_id , type = "response",  
                at = list(rc_trial_time = plannedtimes) )
  cont <- contrast(em, interaction = c("consec", "consec")) 
  CI <- confint(cont) 
  Su <- summary(cont)
  coef <- as.data.frame(cont)$ratio
  confint <- CI[,c("asymp.LCL", "asymp.UCL")]
  
  
  zstat = round(Su$z.ratio,3)
  pz = ifelse (Su$p.value < 0.001, "<0.001",paste("= ",round(Su$p.value,3)))
  unadj_res = cbind(coef = coef, confint, zstat = zstat, pz = pz, likrat)
  
  #adjusted
  
  anova_call <- anova(adj_null,adj_mod)
  likrat <-  data.frame(Chisq = paste0(round(anova_call$Chisq[2],3),"(",
                                       anova_call$Df[2],")"),
                        pvalue = anova_call$`Pr(>Chisq)`[2])
  em <- emmeans(adj_mod,  ~ rc_trial_time*group_type_id , type = "response",  
                at = list(rc_trial_time = plannedtimes) )
  cont <- contrast(em, interaction = c("consec", "consec"))
  CI <- confint(cont) 
  Su <- summary(cont)
  coef <- as.data.frame(cont)$ratio
  confint <- CI[,c("asymp.LCL", "asymp.UCL")]
  
  
  zstat = round(Su$z.ratio,3)
  pz = ifelse (Su$p.value < 0.001, "<0.001",paste("= ",round(Su$p.value,3)))
  adj_res = cbind(coef = coef, confint, zstat = zstat, pz = pz, likrat)
  
  
  # calculate CI
  res_tab <- rbind(adj_res,unadj_res) %>% 
    as.data.frame() %>% 
    remove_rownames %>% 
    mutate(response = params$response,
           adjusted = c("No","Yes")) %>% 
    transmute(response, adjusted,
              planned = coef, asymp.LCL = asymp.LCL, asymp.UCL = asymp.UCL, 
              Chisq, p.value = pvalue)
  
  
  
  # save models
  write.csv(res_tab,file = paste0("../../Results/baseline_sd/",params$response,"_table.csv"), row.names = FALSE)
  

  
} else {
  res_tab <- read.csv(file = paste0("../../Results/baseline_sd/",params$response,"_table.csv"))
  
  
}

res <- res_tab %>% 
  transmute(response,
            adjusted,
            base_3yr_ratio = round(planned,3),
            LCL = round(asymp.LCL,3),
            UCL = round(asymp.UCL,3),
            zstat,
            pz,
            Chisq,
            `p.value` = ifelse (`p.value` < 0.001, "<0.001",paste("= ",round(`p.value`,3)))
)

res


evidence_unadj = ifelse(res_tab$p.value[1]<0.05, "evidence", "no evidence")
evidence_adj = ifelse(res_tab$p.value[2]<0.05, "evidence", "no evidence")




```

```{r, echo=!params$final, message=!params$final, warning=!params$final}

# calculate all ems by time and group
group_time_ems <- emmeans(fitted_mods$unadj_mod,  ~ group_type_id  + rc_trial_time, type = "response")

# contrast to baseline by group
comp_to_baseline <- contrast(group_time_ems, rc_trial_time , "trt.vs.ctrl", by=  "group_type_id")

time_diff <- c(1:3)

to_bl_plot_df <- cbind(summary(comp_to_baseline)[,1:3], 
                       confint(comp_to_baseline)[,6:7], 
                       `Visit (Years)` = rep(c(1:3), 2))
# add baseline 0
to_bl_plot_df[7:8,] <- NA
to_bl_plot_df$group_type_id[7:8] = c(1,2)
to_bl_plot_df$ratio[7:8] = c(1,1)
to_bl_plot_df$`Visit (Years)`[7:8] = 0


to_bl_plot_df %>% 
  ggplot(aes(`Visit (Years)`, ratio, color=factor(group_type_id), group=factor(group_type_id))) + 
  geom_point() + 
  geom_line()+
  geom_linerange(aes(ymin=asymp.LCL, ymax=asymp.UCL), size=2, alpha=.5) + 
  theme_classic() + 
  ylab("Relative change from baseline") +
  scale_color_discrete(name="Randomisation Group")  + 
   theme(legend.position = "bottom") 

ggsave(file = paste0(params$response,"_",params$data,"_plot.png"), width = 12, height = 8, units = "cm")



to_bl_plot_df %>% 
  transmute(group_type_id, `Visit (Years)`,
         ratio,
         LCL = asymp.LCL, UCL = asymp.UCL) %>% 
  arrange(`Visit (Years)`,group_type_id) %>% 
write.csv(file = paste0(params$response,"_",params$data,"_plotdata.csv"), row.names = FALSE)



ratios_adj <- emmeans(fitted_mods$adj_mod, trt.vs.ctrl ~ group_type_id  + rc_trial_time, type = "response", by=  "group_type_id", adjust = NULL) $contrasts[c(3,6)] %>% confint()

ratios_adj[,3:7] <- round(ratios_adj[,3:7],2)

ratios_unadj <- emmeans(fitted_mods$unadj_mod, trt.vs.ctrl ~ group_type_id  + rc_trial_time, type = "response", by=  "group_type_id", adjust = NULL) $contrasts[c(3,6)] %>% confint() 

ratios_unadj[,3:7] <- round(ratios_unadj[,3:7],2)


```


## Result


These data are positively skewed and bounded by 0. As such, a normal disitrubution is not recommended. We tried a number of GLMMs with capacity to handle such distributions, and concluded a tweedie distribution was most appropriate for these data.

. 
Covariates included in adjusted models: 

  - `myb_age`: Age at Baseline 
  - `myb_gender`: Reported gender
  - `cl_edu_yrs`: Years of education
  - `rcal_modules_sum`: Number of modules person was eligible for
  - `adri_baseline`: ADRI at Baseline


A tweedie mixed model with a random effect for individual and fixed effects of time point, randomisation group and their interaction was run. An additional analysis including covariates of `r paste(params$covariates, collapse=", ")` was run.  We report estimate and confidence interval for the relative change in minutes per week within each group, with assoicated z statistic and p-value. 

For group = 1, the unadjusted estimated relative change is `r ratios_unadj$ratio[1]` from baseline to 3 years (95%CI `r ratios_unadj$asymp.LCL[1]` - `r ratios_unadj$asymp.UCL[1]`), for group 2, the estimated relative change is `r ratios_unadj$ratio[2]` from baseline to 3 years (95%CI `r ratios_unadj$asymp.LCL[2]` - `r ratios_unadj$asymp.UCL[2]`). The unadjusted ratio between randomisation groups of relative change from baseline to three year follow up was `r res$base_3yr_ratio[1]` ( 95%CI: `r res$LCL[1]`- `r res$UCL[1]`, Z=`r res$zstat[1]`,p`r res$pz[1]` ). 



For group = 1, the adjusted estimated relative change is `r ratios_adj$ratio[1]` from baseline to 3 years (95%CI `r ratios_adj$asymp.LCL[1]` - `r ratios_adj$asymp.UCL[1]`), for group 2, the estimated relative change is `r ratios_adj$ratio[2]` from baseline to 3 years (95%CI `r ratios_adj$asymp.LCL[2]` - `r ratios_adj$asymp.UCL[2]`). The unadjusted ratio between randomisation groups of relative change from baseline to three year follow up was `r res$base_3yr_ratio[2]` ( 95%CI: `r res$LCL[2]`- `r res$UCL[2]`Z=`r res$zstat[2]`,p`r res$pz[2]` ).   



Analysis by Nancy Briggs and Gordana Popovic on `r Sys.Date()`


