---
title: Maintain your brain, Primary Cognitive Composite Variable sugbroup by Age < 66
output: html_document
params:
  data : "cogprimary"
  response : "cogmeanz_full"
  predictors : !r c("rc_trial_time","group_type_id","age_bin")
  covariates : !r c("rcal_modules_sum", "myb_gender","cl_edu_yrs","adri_baseline")
  id : "id_mask"
  final : true
---

## `r params$response` by subgroups of `r params$predictors[3]`

Analyses for the primary composite outcome, three-way interaction between age (<66 v.s. >=66), group and time.

Age groups have been updated from the previous definition of <65 vs >=65 

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r read-data, message=FALSE, warning=FALSE}
# read in functions and data
source("../../Programs/functions.R")
emm_options(pbkrtest.limit = 1000)
myb_dat <- read.csv(paste0("../../Data/",params$data,".csv")) %>% 
  mutate(age_bin = ifelse(myb_age < 66,"< 66",">= 66"),
         rc_trial_time = factor(rc_trial_time))
```



## Data summary



```{r outcome-summary, message=!params$final, warning=!params$final}
myb_dat %>% 
  group_by(.dots = lapply(params$predictors, as.symbol)) %>% 
  summarize_at(.vars = params$response, .funs = list(Mean = mean, Sd = sd, N_obs = ~length(.)), na.rm=T)
```

Confirming age groups

```{r, echo=F}
myb_dat %>% 
  group_by(group_type_id,age_bin) %>% 
  summarize( Min=min(myb_age), Max=max(myb_age))
```


```{r fit-models, message=!params$final, warning=!params$final}
# if want to run
if(!params$final){
  
  fitted_mods <- NULL
  fitted_mods$unadj_mod <- lmer(cogmeanz_full ~ rc_trial_time * group_type_id * age_bin + 
                                  (1|id_mask), data=myb_dat)
  fitted_mods$unadj_null <- lmer(cogmeanz_full ~ rc_trial_time * group_type_id + age_bin + 
                                  (1|id_mask), data=myb_dat)
  fitted_mods$adj_mod <- lmer(cogmeanz_full ~ rc_trial_time * group_type_id * age_bin + 
                                rcal_modules_sum + myb_gender + cl_edu_yrs + adri_baseline +
                                  (1|id_mask), data=myb_dat)
  fitted_mods$adj_null <- lmer(cogmeanz_full ~ rc_trial_time * group_type_id + age_bin + 
                                rcal_modules_sum + myb_gender + cl_edu_yrs + adri_baseline +
                                  (1|id_mask), data=myb_dat)
  
  save(fitted_mods,file = paste0("../../Results/Prelim/",params$response,"_age"))
  
} else {
  load(file = paste0("../../Results/Prelim/",params$response,"_age"))
}
```



```{r plot-residuals, message=!params$final, warning=!params$final}
#plot models
if(!params$final){
  lapply(fitted_mods, plot, form = cond)
}
```

```{r print-summary, message=!params$final, warning=!params$final}
#check summary is okay
if(!params$final){
  lapply(fitted_mods, summary)
}
```

## Results

```{r inference, message=!params$final, warning=!params$final}
if(!params$final){
  
  # unadjusted
  anova_call <- anova(fitted_mods$unadj_null,fitted_mods$unadj_mod)
  likrat <-  data.frame(Chisq = paste0(round(anova_call$Chisq[2],3),"(",
                                       anova_call$Df[2],")"),
                        pvalue = anova_call$`Pr(>Chisq)`[2])
  
  
  
  # two way int EM
  
  em <- emmeans(fitted_mods$unadj_mod,  ~ rc_trial_time*group_type_id |age_bin,  
                at = list(rc_trial_time = plannedtimes) )
  cont <- contrast(em, interaction = c("consec", "consec"))
  CI <- confint(cont) %>% 
    dplyr::select(age_bin, asymp.LCL, asymp.UCL) 
    Su <- summary(cont)
  cont <- cont %>% as.data.frame() %>% 
    dplyr::select(age_bin, estimate) 
  
  res_unadj <- left_join(cont, CI)
  res_unadj$zstat = round(Su$z.ratio,3)
  res_unadj$pz = ifelse (Su$p.value < 0.001, "<0.001",paste("= ",round(Su$p.value,3)))
  res_unadj$Chisq = c(likrat$Chisq, NA)
  res_unadj$pvalue = c(likrat$pvalue, NA)
  
    # adjusted
  anova_call <- anova(fitted_mods$adj_null,fitted_mods$adj_mod)
  likrat <-  data.frame(Chisq = paste0(round(anova_call$Chisq[2],3),"(",
                                       anova_call$Df[2],")"),
                        pvalue = anova_call$`Pr(>Chisq)`[2])
  em <- emmeans(fitted_mods$adj_mod,  ~ rc_trial_time*group_type_id |age_bin,  
                at = list(rc_trial_time = plannedtimes) )
  cont <- contrast(em, interaction = c("consec", "consec"))
  CI <- confint(cont) %>% 
    dplyr::select(age_bin, asymp.LCL, asymp.UCL) 
  Su <- summary(cont)
  cont <- cont %>% as.data.frame() %>% 
    dplyr::select(age_bin, estimate) 
  
  res_adj <- left_join(cont, CI)
  res_adj$zstat = round(Su$z.ratio,3)
  res_adj$pz = ifelse (Su$p.value < 0.001, "<0.001",paste("= ",round(Su$p.value,3)))
  
  res_adj$Chisq = c(likrat$Chisq, NA)
  res_adj$pvalue = c(likrat$pvalue, NA)
  
  
  res_tab <- rbind(res_unadj, res_adj) %>% 
     as.data.frame() %>% 
    remove_rownames %>% 
    mutate(response = params$response,
           adjusted = c("No","No","Yes","Yes")) %>% 
    transmute(response = response, adjusted, age_bin,
              planned = estimate, asymp.LCL = asymp.LCL, asymp.UCL = asymp.UCL, 
              zstat, pz, Chisq, p.value = pvalue)
  
  
  # save models
  write.csv(res_tab,file = paste0("../../Results/Prelim/",params$response,"_age","_table.csv"), row.names = FALSE)
  
  
} else {
  res_tab <- read.csv(file = paste0("../../Results/Prelim/",params$response,"_age","_table.csv"))
  
  
}
res <- res_tab %>% 
  transmute(response,
            adjusted,
            age = age_bin,
            base_3yr_did = round(planned,3),
            LCL = round(asymp.LCL,3),
            UCL = round(asymp.UCL,3),
            zstat, 
            pz,
            Chisq,
            `p.value` = ifelse (`p.value` < 0.001, "<0.001",paste("= ",round(`p.value`,3)))
)
res
evidence_unadj = ifelse(res_tab$p.value[1]<0.05, "evidence", "no evidence")
evidence_adj = ifelse(res_tab$p.value[2]<0.05, "evidence", "no evidence")
```

```{r, message=F, warning=F}
# calculate all ems by time and group
group_time_ems <- emmeans(fitted_mods$unadj_mod,  ~ group_type_id + rc_trial_time| age_bin)

# contrast to baseline by group
comp_to_baseline <- contrast(group_time_ems, rc_trial_time, "trt.vs.ctrl", by=  c("group_type_id", "age_bin"))

time_diff <- c(1:3)

to_bl_plot_df <- cbind(summary(comp_to_baseline)[,1:4], 
                       confint(comp_to_baseline)[,7:8], 
                       `Visit (Years)` = rep(c(1:3), 2))
# add baseline 0
to_bl_plot_df[13:16,] <- NA
to_bl_plot_df$group_type_id[13:16] = c(1,2,1,2)
to_bl_plot_df$estimate[13:16] = c(0,0,0,0)
to_bl_plot_df$`Visit (Years)`[13:16] = 0
to_bl_plot_df$age_bin[13:16] = c("< 66","< 66",">= 66",">= 66")

to_bl_plot_df %>% 
  ggplot(aes(`Visit (Years)`, estimate,color=factor(group_type_id), group=factor(group_type_id))) + 
  geom_point() + 
  geom_line()+
  geom_linerange(aes(ymin=asymp.LCL, ymax=asymp.UCL), size=2, alpha=.5) + 
  theme_classic() + 
  facet_grid(~ age_bin)+ 
  ylab("Mean change from baseline") +
  scale_color_discrete(name="Randomisation Group")  + 
   theme(legend.position = "bottom") 

ggsave(file = paste0(params$response,"_",params$data,"_subgroupage_plot.png"), width = 12, height = 8, units = "cm")



to_bl_plot_df %>% 
  transmute(group_type_id, `Visit (Years)`,age_bin,
         estimate,
         LCL = asymp.LCL, UCL = asymp.UCL) %>% 
  arrange(`Visit (Years)`,group_type_id) %>% 
write.csv(file = paste0(params$response,"_",params$data,"_subgroupage_plotdata.csv"), row.names = FALSE)

```


## Result

Covariates included in adjusted models: 

  - `myb_gender`: Reported gender
  - `cl_edu_yrs`: Years of education
  - `adri_baseline`: ADRI at Baseline
  - `rcal_modules_sum`: Number of modules person was eligible for



A linear mixed model with a random effect for patient and fixed effects of time point, randomisation group, binary age (<66 vs >= 66) and their interaction was run. An additional analysis including covariates of `r paste(params$covariates, collapse=", ")` was run.

We report p-values for the time point by randomisation group by binary age (<66 vs >= 66) three-way interaction.  We report estimates and confidence intervals for difference between randomisation groups of change from baseline to three year follow up (two-way interactions, as in the primary analysis) for those with age <66 and >=66  separately. The units of the effect size are standard deviations at baseline. 

The unadjusted difference between randomisation groups of change from baseline to three year follow up was `r res$base_3yr_did[1]` ( 95%CI: `r res$LCL[1]`- `r res$UCL[1]`) baseline standard deviations for patients <66 years and `r res$base_3yr_did[2]` ( 95%CI: `r res$LCL[2]`- `r res$UCL[2]`) baseline standard deviations for patients >= 66 years at baseline. We have `r evidence_unadj` that the two-way group*time interaction is different between the age groups (p`r res$p.value[1]`; this is the 3-way interaction p-value.).  

The adjusted difference between randomisation groups of change from baseline to three year follow up was `r res$base_3yr_did[3]` ( 95%CI: `r res$LCL[3]`- `r res$UCL[3]`) baseline standard deviations for patients <66 years and `r res$base_3yr_did[4]` ( 95%CI: `r res$LCL[4]`- `r res$UCL[4]`) baseline standard deviations for patients >= 66 years at baseline , adjusting for `r paste(params$covariates, collapse=", ")`. We have `r evidence_unadj` the two-way group*time interaction is different between the age groups (p`r res$p.value[1]`; this is the 3-way interaction p-value.).  




Analysis by Nancy Briggs and Gordana Popovic on `r Sys.Date()`



